### 2016-12-04 AuthorResponseToReviewers.Rmd
Expanded on color contrast impact on perception a bit. Added Adelson and Fairchild citations.

---
title: "Authors Response to Reviewers"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## *Binning Strategies and Related Loss for Binned Scatterplots of Large Data*

The following document strives to address comments, questions and concerns made by the reviewers on a point-by-point basis. The reviewer statement are included verbatim through *italic text blocks*; our responses are interspersed to address each point more directly. In the revised article that accompanies, the edits are made in red text to make changes more easily identifiable. The revisions will be reverted to black after the revision process.

### Responses to Reviewer 1 Comments/Questions

*The authors discuss different binning strategies and the resulting aggregation- induced errors. Both errors, loss of spatial resolution and loss of frequency information are introduced and formalized. Binning is an important preprocessing step and research in this area is highly relevant. However, there are some major points missing, making this work less convincing.*

*The most crucial point is in my opinion the missing discussion of the influence of different data distributions. In the visualization community exists research on different descriptors for two-dimensional data called Scagnostics. When I started reading this article, I expected a thorough discussion of different data distributions/properties and the influence on different binning strategies and error outcomes.*

- We are very grateful that the review brought the literature on scagnostics to our attention; the work and scagnostic metrics formalized in the works by Wilkerson, Anand, Grosmann and Wills will be valuable to the discussion of binned scatterplots. Specifically scagnostic literature contributes to Section 5 of our paper in the exploration of loss under a variety of simulated distribution. You will find that a discussion of the relevance of scagnostics has been added immediately preceding the start of subsection 5.1. It is interesting that the striation score is so small, given that the data resolution issue can lead to artificially induced stripping is seen in so prominently in Figure 6.

*In Section 4.2, a discussion on contrast effects hindering the perception of the encoded values is missing. Nearby colored points, will change the color impression. These effects can be modelled and estimated and maybe interesting to investigate a little further.*

- Citations to refer readers to Adelson and Fairchild as example of work on contrast effects has been added to Section 4.2 and related future work has been added to the discussion in Section 6.

*As a computer scientist, it is not clear to me, why the runtime is larger with lower numbers of bins (Section 5.1. lines 33ff.)  In my naïve implementation, I would iterate once over all data points and assign the respective bins by an adopted integer division. Consequently, the runtime would be in O(n d) with n being the number of data points and d the number of dimensions. Maybe, the authors could explain the runtime experiments more or discuss the used algorithm more. If the number of bins is 1, I would expect the fastest runtime.*

- Your intuition is correct, but there was a misunderstanding about Figure 5. The x-axis relates the the *size* of bins not the *number* of bins, meaning that the binning takes less time as the bin widths grows (ie the number of bins decreases). x=1 means using many 1X1 bins, not one large bin. The x-axis label has been edited to more clearly reflect that this is the size of the bin dimensions.

*In the application example, why are 50 bins chosen and why is standard rectangular binning chosen? How would other binning results differ from the presented binning. The different variants in Figure 11 are only different color mappings of density. I would include here a discussion of data properties (Scagnostics) together with some error measurements described by the authors beforehand.*

- The beginning of section 5.3 has been rewritten to more explicitly tie the binning choices for the baseball example to the lessons learned in 5.1 and 5.2.

*Minor remarks:Add color legend to Figure 7*

- Added as per request.

### Responses to Reviewer 2 Comments/Questions

*This paper reviews visualization methods for scatterplots and their strengths and weaknesses in handling large datasets. Section 2 notes that binned aggregation is the most widely used and effective way to visualize big bivariate data. In Section 3, the paper introduces three types of binning algorithms: standard rectangular, random and quantile. The main idea in this paper is presented in Section 4, where loss of information is characterized using two metrics: spatial loss and frequency loss. Spatial loss accounts for the loss of information about each individual point’s location; and frequency loss addresses the limitation of human perception in distinguishing frequencies represented using hues or saturation. Section 5 then presents some empirical results on how various factors may affect spatial loss and frequency loss based on three synthetic datasets. These factors include bin width and bin origin for spatial loss, and number of bins for frequency loss.*

*In general, I think this paper addresses an important topic. Guidelines on choosing binning algorithms will be immensely helpful for practitioners dealing with big data. I am, however, reserved in recommending acceptance because of two main concerns: lack of novelty and unconvincing results.*

*Lack of Novelty:*
*Section 2 and 3 do a good job in reviewing the concepts of binned aggregation and varieties of binning algorithms. However these are fairly well known and the discussion does not seem to contribute novel ideas. The only thing that I haven’t seen before is the random binning approach. Is this a new idea proposed by you? If so, please state clearly. I also wonder how and why random binning can be useful. Please provide scenarios in which it would make sense to use random binning. For the other ideas covered in these two sections, please properly cite relevant work.*

- You are correct that many of the ideas in Sections 2 and 3 are not original, with the exception of the random binning approach, where the methodologies for random-binned aggregation and post processing for net spatial loss are original to our best knowledge. The goal of these sections is to bring many binning related concepts under a unified notation for the benefit of clarity in the sections to follow. For the random binning, clearer language of proposal of the idea and a statement has been added to address the motivation for its introduction. For the unoriginal ideas, additional citations have been made to place our unifying notation better within the existing literature.

*Unconvincing Results:*
*In Section 4.1, you propose the idea of spatial loss as a metric to account for the loss of information. Please explain how this approach is related to Scott’s approach of using mean integrated square error. What makes your definition a better choice?*

- The comparison and benefit over Scotts MISE is not included to the discussion leading into Section 4.

*I do not understand how binned frequencies (as compared to true frequencies) are calculated in Section 4.2 on frequency loss. Please provide an example. Also please clarify is binned frequencies are measures of frequencies perceived by a human.*

- A clarifying example using the visual from Figure 3 has been added to Section 4.2.

*The first paragraph of Section 5.1 shows that spatial loss increases as bin size increases, this seems hardly surprising. In the second paragraph, you talk about three properties of visual estimator: unbiasedness, consistency and efficiency, I think these properties are interesting, but the rest of the paper does not address them sufficiently.*

- The interesting parallels with common properties of statistical estimators were used as a motivating concept for the construction of binning loss functions. We felt that an extended exploration of this analog would distract from the exploration on optimizing bin specification. The need for future work on exploring this topic has been added to the discussion in Section 6.

*I also fail to understand how the discussion on artificial striping in Section 5.1 is relevant for spatial loss.*

- Artificial striping is a symptom of poorly aligned bin sizes and the data resolution, which will lead to imbalanced bin allocation and thus more spatial loss. A clarification of this idea and a visual comparison of bin sizes and spatial loss have been added to section 5.1 in the context of Figure 6.

*Please provide quantified results on spatial loss when you talk about how binning origin affects spatial loss in Figure 7.*

- The statement addressing the Figure already contains a quantified comparison of spatial loss under the two scenarios. **"As an example, Figure 7 displays the binned scatterplots of ten-by-ten unit bins for the fine exponential data where the binning origin at (0,0) incurs seven percent lower spatial loss than for the (-9,-9) origin due to the heavy overlap into negative regions"** The difference is given as a percent decrease in spatial loss because the relative spatial loss change seemed most pertinent.

*In visualization practices, it is common to use both ordinal scales as well as continuous scales when mapping numbers to colors. How might you address the issue of frequency loss when continuous scales are used?*

- The need for future perception research to quantify the implicit frequency loss when using continuous shading scales is now noted in the discussion section of the paper.

*I don’t buy the advice to “use between four and seven frequency”. There simply isn’t empirical evidence in the paper to support this claim.*

- We do not want to allow more color bin than we can *easily* differentiate because the our expressed purpose for working with discretized color scales in this scenario is to have made loss of frequency information *entirely* explicit. While working with discretizing shades we found ourselves worrying about our ability to perfectly match bins to a legend after approximately 6 shades. We do not claim to have gathered empirical evidence, rather we defer to Healey and Enn's Theory on discernible colors (typically 4-7) to back out experience. Figure 9 was included for the reader to assess if they felt our suggestion was justifiable. A statement clarifying that the recommendations is to bolster the reliable perceptual mapping has been added to the discussion section.

*From a human-computer interaction perspective, it is also limiting when the paper talks about spatial loss and frequency loss without any mention of user tasks (e.g. comparison, outlier detection). Do these metric matter for all the tasks involved in reading the scatterplots?*

*Summary*
*This paper feels like a survey paper but the references are not properly cited. As a research paper, it is not clear to me what are the main contributions, and how these contributions build on top of prior work (given that there is no “related work” section in the paper). While the two loss metrics seem potentially interesting, I think more empirical studies are needed with real human participants. It will also improve the paper if the authors can provide take-away guidelines or algorithms that can be applied to real-world problems.*

- The main result from the paper is to provide simple adjustment recommendations for binning specification for reducing loss in binned scatterplots. To emphasize these recommendations, the baseball plot example in Section 5.3 has been reworked to highlight the take-home message. To your other point, we agree that future work to quantify the implicit loss of information (e.g. the imprecise ability to map continuously shaded tiles to a frequency legend) in reading binned scatterplots is needed; however, the current work strives to quantify explicit loss of information when the perception tasks are more simple (e.g. mapping distinguishably-discrete shaded tiles to an interval on the frequency legend).

### Responses to Editors Comments/Questions
*Editor's Comments to Author:*

*Associate Editor - Comments to the Author:*
*The paper in its current form does not provide a sufficiently thorough coverage of either review material or novel material to warrant publication.  However, the reviewers have some useful suggestions for ways in which a different paper (or two) could possibly be produced by expanding on different parts of this submission.*

- We hope that the revisions explanding review material and clarification of our contributions to the topic will meet the standards of your publication. If there are any additional comments, concerns or questions, our team is happy to address them.